Differential invariant signatures for planar Lie group transformations of images

### 2020-11-16 paper/newpaper.tex
refs
bibupdates

\documentclass[review,onefignum,onetabnum]{siamonline190516}
\usepackage{amsmath,amsfonts,graphicx}
\def\R{\mathbb{R}}
\usepackage{xcolor}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{array}

@@

For image recognition it is then natural to seek representations of objects that do not change as these various transformation groups act. If the space of objects is denoted $M$ and the appearance transformations form a group $\mathcal{G}$, then mathematically one seeks a way to study objects in $M / \mathcal{G}$. One way to achieve this is by transforming all objects to match a chosen target object by registration (the transformation of the appearance of one shape to match another, typically by gradient descent on a matching function, see e.g.~\cite{}). Hence for the group of similarity transformations (rotation and reflection, translation, and scale), a Procrustes alignment~\cite{} will place all of the shapes into the reference frame of one template example. However, analysis on these quotient spaces is rarely simple~\cite{}, and if one is interested in the object, and not the transformation, then it necessitates a lot of computational work for relatively little benefit.

An alternative option is to seek $\mathcal{G}$-invariant functions on $M$, which is what we consider here. In this case, rather than considering the image of the object, one works with an invariant signature of some kind, which represents the object modulo the action of elements from the relevant group. These signatures are typically of a higher dimensionality that the original image.

The theory of invariants has a rich mathematical history~\cite{}.

Olver et al. joint differential invars (aka semi-diff)
MOMENTS

In order for the invariant signatures to be useful to identify similarity in images, there are two key considerations: they must vary continuously with respect to the image, either for all smooth images or for all except a set of sufficiently high codimension, and they must be robust with respect to occlusions, where one object obscures some part of another. These constraints lead to the 
This leads one naturally to consider local invariants, i.e., those defined using each point of the object as its own basepoint, together with a neighbourhood in some topology. This is sufficient to compute, e.g., derivatives of an object. 
This fails for a variety of 
However, of particular saliance for object recognition in scenes is the need to be robust to occlusions,

The introduction of an invariant signature reduces the recognition problem to one of comparing the higher-dimensional signatures that encode the desired self-equivalences. In this paper we introduce the mathematical considerations necessary to construct invariant signatures to the planar Lie groups that can transform 2D images. We consider a wide range of planar groups, both finite-dimensional Lie groups and infinite-dimensional pseudo-Lie ones; as we shall show, the lattice structure of the set of groups influences the relationship between the sets of invariants for the different groups. We provide sample signature sets for the various groups we consider, and examples of the appearance of those signatures for transformations of a simple, smooth image, but do not consider the question of a complete practical implementation. We do not consider the matching of the signatures, for which several methods already exist, such as~\cite{}.

derivatives: noise is not
not intensity transforms

@@

In order to define differential invariants, it is necessary to identify how the group transformation $\varphi$ acts on derivatives of the function $f$. This is known as the jet space, or jet bundle~\cite{OlverJets}. The $d$-th order jet space $J^d$ consists of the function itself and the set of all partial derivatives of the function up to order $d$. (Note that there are $\left( \begin{array}{c} d+k-1 \\ k \end{array} \right)$ partial derivatives of order $d$). As the group $\mathcal{G}$ is local, it acts naturally on derivatives of the transformation, and so each individual transformation $\varphi \in \mathcal{G}$ can be prolonged to act on the jet space $J^d$. We can thus prolong the action of the group and consider the action of $\mathcal{G}$ on not just $f$, but all derivatives of $f$ up to order $d$. A $d$-th order differential invariant is then a local scalar function $I: J^{d} (f,\Omega, k) \to \mathbb{R}$ that is invariant under the action of $\mathcal{G}$, i.e., $I(\varphi(f,\Omega, k)) = I(f, \Omega, k) \forall \varphi \in \mathcal{G}$.

@@

\subsection{Relevant Literature}
---
TEXT FROM OUR MOBIUS PAPER
Calabi et al. [9] propose the use of differential invariant signatures for shape analysis, and further
argue that these should be approximated in a group-invariant way. For example, for the Euclidean
group, the signature is the shape \todo[inline]{something didn't copy paste correctly here} (??; ??s) regarded as a subset of R2. The claimed advantages
of the approach are that the signature determines the shape; that it does not depend on the
choice of initial point on the curve or on parameterization by arclength, and that its G-invariance
makes it robust; and that it is based on a general procedure for arbitrary objects and groups.
See [2, 4, 9, 18, 30, 35, 36] for further developments and applications of the invariant signature.
Although the signature at first sight appears to be complete (e.g., Theorem 5.2 in [9]),
a more detailed treatment (e.g., [17, 28]) highlights the fact that it is not complete on shapes
that contain singular parts { straight and circular segments in the Euclidean case. For these
parts, the signature reduces to a point. Thus, for shapes that have nearly straight or nearly
circular segments, the signature cannot be robust. In addition, while the signature does not
depend on the starting point or the parameterization, it takes values in a very complicated set,
namely, the planar shapes. To compare the invariants of two shapes requires comparing two
shapes. Essentially, the parameterization-dependence has only been deferred to a later stage
of the analysis (unless one is content to compare shapes visually). One approach to this is to
weight the signature, see [32] for more details.
---

@@

An algorithm to compute invariants and determine complete signature sets is
described in \cite{olver2001}. First, the entire set of invariants is
computed using the method of moving frames. Second, determine the
differential invariant order $s$ of $f$ \todo[inline]{we need to define the differential invariant order}. Then the signature set of a maximal
set of functionally independent invariants of order $s+1$ is complete. A
third step is possible, in which the dimension of the signature set is
reduced by eliminating those invariants whose values can be computed from the
values of derivatives of some subset of the invariants.

@@

We will demonstrate this for the affine group $A(2)$ and its subgroup
$SA(2)$ where, following the treatment in~\cite{OlverCIT}, this leads to
Cayley's Omega Process~\cite{Cayley}.

@@

This can be generalised to $n$ functions $f^{(1)}(x_1, y_1), \ldots, f^{(n)}(x_n, y_n)$ and $r$-th order~\cite{Olver} to define a partial transvectant as:

@@

A moving frame is a mapping into a group $\mathcal{G}$ that provides a frame of reference along a curve. Cartan's method of moving frames generalises the concept to a homogeneous space of a Lie group $\mathcal{G}$. When defined on the principal bundle it defines a $\mathcal{G}$-equivariant mapping, which enables one to identify invariants under the action of groups of transformations. The method of moving frames has a long history, as summarised in \cite{Olver14}; particularly noteworthy contributions are by Cartan~\cite{Cartan37} and Olver and co-authors; see~\cite{Olver14} for an overview, and note particularly for an application to~\cite{Calabi98}. The method is described in detail in~\cite{OlverCIT}.

@@

These invariants are a functionally independent, complete, set of polynomial differential invariants up to second derivatives. Note that the invariants found previously by tensor contraction (\cite{eq:se2}) can be expressed in terms of these:
\begin{align*}
  f = f &= I_0 \\
  f_if_i = f_x^2 + f_y^2 &= I_1 \\
  f_{ii} = f_{xx}^2 + f_{yy}^2 &= \frac{I_2 + I_4}{I_1} \\
  f_{ij} f_{ij} = f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2 &= I_2\\
  f_i f_j f_{ij} = f_x^2 f_{xx} + 2f_x f_y f_{xy} + f_y^2 f_{yy} &= \frac{I_2^2 + I_4^2 + 2I_3^2}{I_1^2}.
\end{align*}

@@

Identifying $\R^2$ with $\mathbb{C}$, the M\"obius group is a real 6-dimensional local Lie group acting on $\mathbb{R}^2$.
It is the smallest nonlinear planar group that contains $SE(2)$, and it also has direct applications in image processing since it arises in the {\em conformal camera} model of vision, in which scenes are projected radially onto a sphere \cite{lenz1990group,turski2004geometric}; it is also the set of biholomorphic maps of the Riemann sphere. 
%Different shapes may be related by M\"obius transformations, as is explored by Petukhov \cite{petukhov1989non} for biological objects: in one example, the human skull grows approximately by M\"obius transformations.  Note also that Thompson \cite{thompson1942growth} famously deformed images of one species to match those of another, many of his examples using conformal mappings, and thus may be approximated (or even determined by) M\"obius transformations. Discussing this work, Milnor \cite{milnor2010growth}  argued that ``The geometrically simplest way to change the relative size of different body parts would be by a conformal transformation. It seems plausible that this simplest solution will often be the most efficient, so that natural selection tend to choose it.'' (Milnor was thinking of 3D conformal transformations, whose restriction to 2D is the M\"obius transformations.)

@@

\todo[Move this one out]

@@

>>>>

\documentclass[review,onefignum,onetabnum]{siamonline190516}
\usepackage{amsmath,amsfonts,graphicx}
\def\R{\mathbb{R}}
\usepackage{xcolor}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{array}
\usepackage{natbib}

@@

For image recognition it is then natural to seek representations of objects that do not change as these various transformation groups act. If the space of objects is denoted $M$ and the appearance transformations form a group $\mathcal{G}$, then mathematically one seeks a way to study objects in $M / \mathcal{G}$. One way to achieve this is by transforming all objects to match a chosen target object by registration (the transformation of the appearance of one shape to match another, typically by gradient descent on a matching function, see e.g.~\cite{Modersitzki03}). Hence for the group of similarity transformations (rotation and reflection, translation, and scale), a Procrustes alignment~\cite{Kendall1989} will place all of the shapes into the reference frame of one template example. However, analysis on these quotient spaces is rarely simple (see e.g.,~\cite{Marsland19b} for a basic primer), and if one is interested in the object, and not the transformation, then it necessitates a lot of computational work for relatively little benefit.

An alternative option is to seek $\mathcal{G}$-invariant functions on $M$, which is what we consider here. In this case, rather than considering the image of the object, one constructs an invariant signature of some kind, which represents the object modulo the action of elements from the relevant group. This has been considered extensively by Olver and co-authors for the case of outlines of objects~\cite{Calabi1998,Hoff2013}. In this paper we consider the case of signatures for images rather than curve outlines.

The introduction of an invariant signature reduces the recognition problem to one of comparing the higher-dimensional signatures (typically the signatures need to be at least three dimensional for planar images) that encode the desired self-equivalences. In this paper we introduce the mathematical considerations necessary to construct invariant signatures to the planar Lie groups (and some infinite-dimensional, pseudo-Lie groups too) that can transform 2D images. As we shall show, the lattice structure of the set of groups influences the relationship between the sets of invariants for the different groups. We provide sample signature sets for the various groups we consider, and examples of the appearance of those signatures for transformations of a simple, smooth image, but do not consider the question of a complete practical implementation. We do not consider the classification of the signatures, for which either a standard image recognition algorithm can be used (see, e.g.,~\cite{Zhang20} for a survey), or a matching method such as~\cite{UsCurrents} can be used.

In order for the invariant signatures to be useful to identify similarity in images, there are two key considerations: they must vary continuously with respect to the image, either for all smooth images or for all except a set of sufficiently high codimension, and they must be robust with respect to occlusions, where one object obscures some part of another. These constraints lead one naturally to consider local invariants, i.e., those defined using each point of the object as its own basepoint, together with a neighbourhood in some topology. This is sufficient to compute, e.g., derivatives of an object. For this reason we focus on differential invariants. The disadvantage of such invariants, as we shall discuss, is the need to compute several orders of spatial derivative of the image.

@@

In order to define differential invariants, it is necessary to identify how the group transformation $\varphi$ acts on derivatives of the function $f$. This is known as the jet space, or jet bundle~\cite{OlverEIS}. The $d$-th order jet space $J^d$ consists of the function itself and the set of all partial derivatives of the function up to order $d$. (Note that there are $\left( \begin{array}{c} d+k-1 \\ k \end{array} \right)$ partial derivatives of order $d$). As the group $\mathcal{G}$ is local, it acts naturally on derivatives of the transformation, and so each individual transformation $\varphi \in \mathcal{G}$ can be prolonged to act on the jet space $J^d$. We can thus prolong the action of the group and consider the action of $\mathcal{G}$ on not just $f$, but all derivatives of $f$ up to order $d$. A $d$-th order differential invariant is then a local scalar function $I: J^{d} (f,\Omega, k) \to \mathbb{R}$ that is invariant under the action of $\mathcal{G}$, i.e., $I(\varphi(f,\Omega, k)) = I(f, \Omega, k) \forall \varphi \in \mathcal{G}$.

@@

\subsection{Relevant Literature}
The theory of invariants has a rich mathematical history, see e.g.,~\cite{OlverCIT,Kraft2002} for an overview. Of most relevance to us here 
\cite{Cayley,Littlewood44,Cartan52,Patterson1928,Veblen1872}

\cite{Olver2005,Cartan35}

not intensity transforms

Signature

Relative invars
Joint sigs
\cite{Fels1997,Olver2013,Arora2009}

Diff Invs
\cite{Florack1993,Olver2007,Bouton1898}

Projective invars
\cite{Suk2000,Suk2004,Hann2002a,Kogan2014}

Moments
\cite{Hickman2011,Abu-Mostafa1984,Kakarala2010,Flusser2009,Papakostas2013}

FFT
\cite{Reddy1996,Gauthier2008,Kakarala2011,Smach2007}

Integral invar
\cite{Manay2006,Negrinho2013,Ghorbel1994,Turski2006}

There has been a long-standing interest in constructing invariants for image recognition.
\cite{VanGool1995,Wood1996,Zhang2019,Faugeras1994,Gros1992,Taubin1992,Lenz1990,Florack1993}

\cite{SURF}

---
TEXT FROM OUR MOBIUS PAPER
Calabi et al. [9] propose the use of differential invariant signatures for shape analysis, and further
argue that these should be approximated in a group-invariant way. For example, for the Euclidean
group, the signature is the shape \todo[inline]{something didn't copy paste correctly here} (??; ??s) regarded as a subset of R2. The claimed advantages
of the approach are that the signature determines the shape; that it does not depend on the
choice of initial point on the curve or on parameterization by arclength, and that its G-invariance
makes it robust; and that it is based on a general procedure for arbitrary objects and groups.
See [2, 4, 9, 18, 30, 35, 36] for further developments and applications of the invariant signature.
Although the signature at first sight appears to be complete (e.g., Theorem 5.2 in [9]),
a more detailed treatment (e.g., [17, 28]) highlights the fact that it is not complete on shapes
that contain singular parts { straight and circular segments in the Euclidean case. For these
parts, the signature reduces to a point. Thus, for shapes that have nearly straight or nearly
circular segments, the signature cannot be robust. In addition, while the signature does not
depend on the starting point or the parameterization, it takes values in a very complicated set,
namely, the planar shapes. To compare the invariants of two shapes requires comparing two
shapes. Essentially, the parameterization-dependence has only been deferred to a later stage
of the analysis (unless one is content to compare shapes visually). One approach to this is to
weight the signature, see [32] for more details.
---

@@

An algorithm to compute invariants and determine complete signature sets is
described in \cite{Olver2005}. First, the entire set of invariants is
computed using the method of moving frames. Second, determine the
differential invariant order $s$ of $f$ \todo[inline]{we need to define the differential invariant order}. Then the signature set of a maximal
set of functionally independent invariants of order $s+1$ is complete. A
third step is possible, in which the dimension of the signature set is
reduced by eliminating those invariants whose values can be computed from the
values of derivatives of some subset of the invariants.

@@

We will demonstrate this for the affine group $A(2)$ and its subgroup
$SA(2)$ where, following the treatment in~\cite{OlverCIT}, this leads to
Cayley's Omega Process~\cite{Cayley1846}.

@@

This can be generalised to $n$ functions $f^{(1)}(x_1, y_1), \ldots, f^{(n)}(x_n, y_n)$ and $r$-th order~\cite{OlverEIS} to define a partial transvectant as:

@@

A moving frame is a mapping into a group $\mathcal{G}$ that provides a frame of reference along a curve. Cartan's method of moving frames generalises the concept to a homogeneous space of a Lie group $\mathcal{G}$. When defined on the principal bundle it defines a $\mathcal{G}$-equivariant mapping, which enables one to identify invariants under the action of groups of transformations. The method of moving frames, which originated with Cartan~\cite{Cartan35}  has a long history, as summarised in \cite{Olver2005}. Olver and co-authors, in particular, have popularised the approach and made it applicable to image analysis, see~\cite{Olver2005} for an overview, and the original application paper ~\cite{Calabi1998}. The method is described in detail in~\cite{OlverCIT}.

@@

These invariants are a functionally independent, complete, set of polynomial differential invariants up to second derivatives. Note that the invariants found previously by tensor contraction (\ref{eq:se2}) can be expressed in terms of these:
\begin{align*}
  f = f &= I_0 \\
  f_if_i = f_x^2 + f_y^2 &= I_1 \\
  f_{ii} = f_{xx}^2 + f_{yy}^2 &= \frac{I_2 + I_4}{I_1} \\
  f_{ij} f_{ij} = f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2 &= I_2\\
  f_i f_j f_{ij} = f_x^2 f_{xx} + 2f_x f_y f_{xy} + f_y^2 f_{yy} &= \frac{I_2^2 + I_4^2 + 2I_3^2}{I_1^2}.
\end{align*}

@@

Identifying $\R^2$ with $\mathbb{C}$, the M\"obius group is a real 6-dimensional local Lie group acting on $\mathbb{R}^2$.
It is the smallest nonlinear planar group that contains $SE(2)$, and it also has direct applications in image processing since it arises in the {\em conformal camera} model of vision, in which scenes are projected radially onto a sphere \cite{Lenz1990,Turski2004}; it is also the set of biholomorphic maps of the Riemann sphere. 
%Different shapes may be related by M\"obius transformations, as is explored by Petukhov \cite{petukhov1989non} for biological objects: in one example, the human skull grows approximately by M\"obius transformations.  Note also that Thompson \cite{thompson1942growth} famously deformed images of one species to match those of another, many of his examples using conformal mappings, and thus may be approximated (or even determined by) M\"obius transformations. Discussing this work, Milnor \cite{milnor2010growth}  argued that ``The geometrically simplest way to change the relative size of different body parts would be by a conformal transformation. It seems plausible that this simplest solution will often be the most efficient, so that natural selection tend to choose it.'' (Milnor was thinking of 3D conformal transformations, whose restriction to 2D is the M\"obius transformations.)

@@

\todo{Move this one out}

@@

\bibliographystyle{plainnat}
\bibliography{invariants}
